

```{r libraries}
library(dplyr)
library(SignalDetectionTool)
```

```{r parameters}
if (is.null(params$data)) {
  data <- SignalDetectionTool::input_example
}
```


# Data and Methods 

- Method used: `r names(available_algorithms()[which(available_algorithms() == params$method)])`
- Length of signal detection period: `r params$number_of_weeks` weeks
- Strata selected: `r paste(params$strata, collapse = ", ")`
- Intervention date: `r params$intervention_date`
`r ifelse(grepl("glm|farrington", params$method) && !is.null(params$alpha_upper), paste0("- P-value cutoff: ", params$alpha_upper), paste0(""))`
- Filters applied in the application: `r paste0(ifelse(length(params$selected_filter_vars) != 0, paste(params$selected_filter_vars, collapse = ", "), "None"))`
`r ifelse(params$exclude_outbreak_cases_from_fitting == TRUE && grepl("glm", params$method), paste0("- Only cases not associated with outbreaks are used for training"), paste0(""))`


```{r}
if (!is.null(params$signals_padded) & !is.null(params$signals_agg)) {
  signals_agg <- params$signals_agg
  signals_padded <- params$signals_padded
} else {
  data_preprocessed <- SignalDetectionTool::preprocess_data(data)
  signal_results <- SignalDetectionTool::get_signals(data_preprocessed,
    stratification = NULL,
    method = params$method,
    alpha_upper = params$alpha_upper,
    number_of_weeks = params$number_of_weeks,
    intervention_date = params$intervention_date,
    exclude_outbreak_cases_from_fitting = params$exclude_outbreak_cases_from_fitting
  )

  if (!is.null(params$strata)) {
    signal_results_stratified <- SignalDetectionTool::get_signals(
      data_preprocessed,
      stratification = params$strata,
      method = params$method,
      alpha_upper = params$alpha_upper,
      number_of_weeks = params$number_of_weeks,
      intervention_date = params$intervention_date,
      exclude_outbreak_cases_from_fitting = params$exclude_outbreak_cases_from_fitting
    )

    signal_results <- signal_results %>%
      dplyr::bind_rows(signal_results_stratified)

    signals_agg <- SignalDetectionTool::aggregate_signals(signal_results_stratified,
      number_of_weeks = params$number_of_weeks
    )
  }

  if (grepl("glm", params$method)) {
    signals_padded <- signal_results
  } else {
    signals_padded <- SignalDetectionTool::pad_signals(
      data_preprocessed,
      signal_results
    )
  }
}
```


# Signal Detection 

```{r, results='asis'}
results_analysis <- signals_padded %>%
  dplyr::filter(
    is.na(category),
    !is.na(alarms)
  )


stringr::str_glue(
  "In the signal detection period - last {params$number_of_weeks} weeks available in the data - ",
  "there were {sum(results_analysis$cases)} cases of {params$disease} in {params$country}. ",
  "Using the {names(params$method)} algorithm {sum(results_analysis$alarms)} signal{ifelse(sum(results_analysis$n_alarms) == 1, '', 's')} {ifelse(sum(results_analysis$n_alarms) == 1, ' was', ' were')} detected."
)
```


```{r}
#| fig.cap=paste0("Number of ", params$disease, " cases by week and signals, ", params$country),
#| out.width="100%"
SignalDetectionTool::plot_time_series(
  signals_padded %>%
    dplyr::filter(is.na(category)),
  intervention_date = params$intervention_date,
  interactive = FALSE
)
```

## Signal Detection Table

```{r}
txt <- NULL

if (sum(results_analysis$alarms, na.rm = TRUE) != 0) {
  SignalDetectionTool::build_signals_table(
    signals_padded %>%
      dplyr::filter(is.na(category)),
    format = "Flextable"
  )
} else {
  txt <- "No signals were detected."
}
```

```{r, results='asis'}
#|eval = is.character(txt)

cat(txt)
```
