---
title: "SignalDetectionTool: R Package for signal detection on infectious disease surveillance data"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{SignalDetectionTool}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.dim = c(4, 3),
  fig.align = "center"
)
```

```{r setup, include=FALSE}
options(digits = 3)

library(dplyr)
```

The **SignalDetectionTool** (SDT) is an R package which contains an R shiny app to perform signal detection on infectious disease surveillance data and visualise results. Furthermore the functions of this R package can also be used standalone without running the shiny application. This vignette gives guidance on how to use the functions within the R package. 

## Running the App 

This section is extensively described in the README of the package thus we give minimal instructions here.
First install and load the SDT package.

```{r}
library(SignalDetectionTool)
```

Run the app entering the following command in the console. 

```{r, eval=FALSE}
run_app()
```

It is possible to run the app with customised settings using a yaml file. Please have a look at the README.

## Using functions from the SDT

### Data
The input to the SDT is a linelist of infectious disease cases structured according to the format defined in `input_metadata`.
A sample linelist (`input_example`) and its corresponding metadata specification (`input_metadata`) is available in the package as internal datasets and can be accessed using:
```{r, eval=FALSE}
data("input_example")
data("input_metadata")
```

### Applying signal detection

To apply signal detection to your data for the most recent 6 weeks you first need to preprocess your linelist and then apply `get_signals()`. Inside `get_signals()` the linelist is aggregated to a timeseries containing weekly counts of cases.  
To apply different signal detection methods use the parameter method. To generate stratified signals use the parameter stratification. Please have a look at the function documentation for more details. 
This is an example for signals generated using the default parameters with FarringtonFlexible and no stratification: 
```{r}
data_prepro <- input_example %>% preprocess_data()
signals <- data_prepro %>% get_signals()
```

The generated `signals` output is a data.frame containing the aggregated number of cases per week with additional columns `cases_in_outbreak`, `alarms`, `upperbound`, `expected`, `category`, `stratum`, `method` and `number_of_weeks`.  

* `cases_in_outbreak` is only generated if the column `outbreak_status` was available in your provided linelist. It counts how many cases in this week were already part of a known outbreak.  
* `alarms` is a column with booleans indicating for which weeks a signal has been generated. Of course the majority of weeks in the aggregated timeseries contain `NA` in this column as only the most recent selected `number_of_weeks` are filled with `TRUE` of `FALSE`.  
* `upperbound` contains a numeric threshold calculated with the signal detection methods. As for `alarms`  the majority of weeks in the aggregated timeseries contain `NA` as this is only calculated for the `number_of_weeks` for which signals are generated. Furthermore some algorithms such as EARS do not calculate a threshold.  
* `expected` is a numeric value giving the expected number of cases by the signal detection algorithm. Only filled for the weeks for which signals are generated.  
* `category` Character string specifying to which stratum timeseries belongs to when having added stratification parameters (more details in the next example). For unstratified timeseries it is `NA`.  
* `stratum` Character string specifying to which category of the stratum the timeseries belongs to when having added stratification parameters (more details in the next example). For unstratified timeseries it is `NA`.  
* `method` Character string specifying the signal detection method which was used. 
* `number_of_weeks` Integer specifying the number of weeks for which signals were generated.  



Generating stratified signals using EARS:
```{r}
data_prepro <- input_example %>% preprocess_data()
signals_ears_stratified <- data_prepro %>% get_signals(
  method = "ears",
  stratification = c("age_group", "county")
)
```

The `signals_ears_stratified` output is a data frame in long format, containing multiple time series of weekly aggregated case counts. Each time series corresponds to a unique combination of the specified stratification variables, with all strata stacked vertically.  

The columns `category` and `stratum` identify the individual timeseries. In this example we obtain the following values:  
* `category` is filled with the column names provided in the `stratification` parameter.  
* `stratum` with the values of these variables.  

In this example we obtain:  
* `category` is filled with "age_group" and "county"  
* `stratum` contains "00–04", "05–09", "10–14", "15–19", "20–24", "25–29", "30–34", "35–39", "40–44",
"45–49", "50–54", "55–59", "60–64", "65–69", "70–74", "75–79", "80–84", "85–89",
"90–94", "95–99", "100–104", "105–109" for rows with `category == "age_group"`.  For rows with `category == "county"` the values of `stratum` are "Burgenland", "Kärnten", "Niederösterreich", "Oberösterreich", "Salzburg", 
"Steiermark", "Tirol", "Vorarlberg", "Wien".

### Create visualisations and tables

To visualise signal detection in your data for the most recent `number_of_weeks`, you first need to preprocess your linelist and then apply `get_signals()`. For some visualisations you then have to aggregate data and need to specify a shapefile to use. In the final step you can apply a function to either create a plot or a table.

The following example shows how to visualise the timeseries of detected signals using the default parameters of FarringtonFlexible without stratification.`plot_time_series()`:
```{r}
data_prepro <- input_example %>% preprocess_data()
signals <- data_prepro %>% get_signals()
time_series <- signals %>% plot_time_series()
```

In the timeseries visualisation, the detection period is shaded in blue. The expectation is shown as a black line. The threshold is indicated by a blue line. Weekly case counts are displayed as bars. A signal is generated for a given week when the number of cases exceeds the threshold.

To create a table with different formatting options you can use `build_signals_table()`. This is an example for creating a DataTable using age_group as `stratum` and default options: 
```{r}
data_prepro <- input_example %>% preprocess_data()
signals <- data_prepro %>% get_signals(stratification = c("age_group"), 
                                       number_of_weeks = 6
                                       )
build_signals_table(signals)
```

To create a non-interactive map you have to use aggregated signals and an appropriate shapefile. Both have to be specified in `create_map_or_table()` and interactive should be set to FALSE. `plot_regional()` is used by the function. This is an example for visualising the specified data using "county" as `stratum` as a map because the matching of the NUTS IDs is fully possible. Otherwise a table would be created. 
```{r}
data_prepro <- input_example %>% preprocess_data() 
signals <- data_prepro %>% get_signals(stratification = c("county"))
signals_agg <- signals %>% aggregate_signals(number_of_weeks = 6)
create_map_or_table(signals_agg, 
                    input_example, 
                    "county", 
                    nuts_shp, 
                    interactive = FALSE
                    )
```

In the non-interactive map visualisation, the number of cases per region is represented by the shade of blue, with darker shades indicating higher case counts. Regions without signals are shown with a black border. If at least one signal occurs in a region, the border is coloured red and the number of signals is displayed as a label within the region.

To create an interactive or a non-interactive barplot or a table you have to use aggregated signals in `create_barplot_or_table()`. Depending on the number of unique levels to visualise it is decided whether a barplot or a table is shown. If you want to visualise a barplot you can use `plot_barchart()` following the same steps. This is an example for visualising the specified data using "age group" as `stratum` and for the non-interactive plot using `create_barplot_or_table()`:
```{r}
data_prepro <- input_example %>% preprocess_data() 
signals <- data_prepro %>% get_signals(stratification = c("age_group"))
signals_agg <- signals %>% aggregate_signals(number_of_weeks = 6)

# Choose one of the two functions: 
create_barplot_or_table(signals_agg, "age_group", 
                        interactive = FALSE
                        )
plot_barchart(signals_agg)
```

In the non-interactive barplot visualisation, the number of cases is represented by the height of the bars. If no signal occurs in a given stratum, the bar has no border. If at least one signal is detected, the bar is outlined in red. For non-interactive plots, the number of signals is additionally displayed as a label above the bar.

To create an interactive or a non-interactive 100% stacked barplot showing a week-to-week representation of strata levels that had alarms you have to use signal data and specify the number of stratification levels in `plot_signals_per_week()`. This is an example for visualising the specified data using "county" as `stratum` with 9 levels for the non-interactive plot:
```{r}
data_prepro <- preprocess_data(input_example)
signals <- get_signals(data_prepro, 
                       stratification = "county"
                       )
n.strata <- 9
plot_signals_per_week(signals, 
                      n_strata = n.strata
                      )
```

In the non-interactive 100% stacked barplot visualisation, the proportion of levels with alarms per week is represented by the height of the red segment of each bar. If no signal occurs in a given week, the bar is shown entirely in grey.

To create a bar plot grouping the "age_group" by another `stratum` use preprocessed data in `plot_agegroup_by()`. This is an example using the `stratum` "county": 
```{r}
data_prepro <- input_example %>% preprocess_data()
plot_agegroup_by(df = data_prepro,
                 age_group_col = "age_group", 
                 by_col = "county"
                 )
```

In the bar plot the height of a bar represents the case count for the specified "county" and "age group". 

### Run the report
#### Single-pathogen reporting

You can directly generate an HTML or Word report containing signal results and visualizations using `run_report()`. Simply pass your surveillance linelist—structured according to the format defined in `input_metadata`—to `run_report()`. The data will be preprocessed automatically within the function. The following code chunk generates a Word report using EARS without any stratification (default) for the past six weeks (default):
```{r, eval=FALSE}
run_report(input_example,
  report_format = "DOCX",
  method = "EARS"
)
```

This generates an HTML report (default) with stratification by age group, county and sex for the last 2 weeks using FarringtonFlexible (default):
```{r, eval=FALSE}
run_report(input_example,
  strata = c("age_group", "county", "sex"),
  number_of_weeks = 2
)
```

#### Multi-pathogen reporting (HTML)

The following functionalities are currently only available for HTML reports. They are not supported in Word (DOCX) format.
The package also supports generating a single HTML report for multiple pathogens.  
To use this functionality, you must provide a line list that includes cases from different pathogens.

An example line list is available in the package as an internal dataset: `input_example_multipathogen`.

You can access it with:

```{r, eval=FALSE}
data("input_example_multipathogen")
```

To run an HTML report with the EARS algorithm and stratification by age_group and county for all pathogens in the line list you can run: 
```{r, eval=FALSE}
run_report(input_example_multipathogen,
  strata = c("age_group", "county"),
  method = "EARS"
)
```

If you want to generate a report for a subset of pathogens in your linelist you can specify these using the pathogens parameter:

```{r, eval=FALSE}
run_report(input_example_multipathogen,
  pathogens = c("Enterobacter", "Salmonella"),
  strata = c("age_group", "county"),
  method = "EARS"
)
```

In addition you can customise the HTML report by replacing the default United4Surveillance logo in the top right corner with your own logo. To do this, provide the path to a .png or .svg file using the custom_logo parameter.

```{r, eval=FALSE}
run_report(input_example_multipathogen,
  pathogens = c("Enterobacter", "Salmonella"),
  strata = c("age_group", "county"),
  method = "EARS",
  custom_logo = "path/to/my/custom_logo.png"
)
```
