---
title: "SignalDetectionTool: R Package for signal detection on infectious disease surveillance data"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{SignalDetectionTool}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.height = 3,
  fig.width = 4,
  fig.align = "center"
)
```

```{r setup, include=FALSE}
options(digits = 3)

library(dplyr)
```

The **SignalDetectionTool** (SDT) is an R package which contains an R shiny app to perform signal detection on infectious disease surveillance data and visualise results. Furthermore the functions of this R package can also be used standalone without running the shiny application. This vignette gives guidance on how to use the functions within the R package. 

## Running the App 

This section is extensively described in the README of the package thus we give minimal instructions here.
First install and load the SDT package.

```{r}
library(SignalDetectionTool)
```

Run the app entering the following command in the console. 

```{r, eval=FALSE}
run_app()
```

It is possible to run the app with customised settings using a yaml file. Please have a look at the README.

## Using functions from the SDT

### Data
The input to the SDT is a linelist of infectious disease cases structured according to the format defined in `input_metadata`.
A sample linelist (`input_example`) and its corresponding metadata specification (`input_metadata`) is available in the package as internal datasets and can be accessed using:

```{r, eval=FALSE}
data("input_example")
data("input_metadata")
```

### Applying signal detection

To apply signal detection to your data for the most recent 6 weeks you first need to preprocess your linelist and then apply `get_signals()`. Inside `get_signals()` the linelist is aggregated to a timeseries containing weekly counts of cases.  
To apply different signal detection methods use the parameter method. To generate stratified signals use the parameter stratification. Please have a look at the function documentation for more details. 
This is an example for signals generated using the default parameters with FarringtonFlexible and no stratification: 

```{r}
data_prepro <- input_example %>% preprocess_data()
signals <- data_prepro %>% get_signals()
```

The generated `signals` output is a data.frame containing the aggregated number of cases per week with additional columns `cases_in_outbreak`, `alarms`, `upperbound`, `expected`, `category`, `stratum`, `method` and `number_of_weeks`.  

* `cases_in_outbreak` is only generated if the column `outbreak_status` was available in your provided linelist. It counts how many cases in this week were already part of a known outbreak.  
* `alarms` is a column with booleans indicating for which weeks a signal has been generated. Of course the majority of weeks in the aggregated timeseries contain `NA` in this column as only the most recent selected `number_of_weeks` are filled with `TRUE` of `FALSE`.  
* `upperbound` contains a numeric threshold calculated with the signal detection methods. As for `alarms`  the majority of weeks in the aggregated timeseries contain `NA` as this is only calculated for the `number_of_weeks` for which signals are generated. Furthermore some algorithms such as EARS do not calculate a threshold.  
* `expected` is a numeric value giving the expected number of cases by the signal detection algorithm. Only filled for the weeks for which signals are generated.  
* `category` Character string specifying to which stratum timeseries belongs to when having added stratification parameters (more details in the next example). For unstratified timeseries it is `NA`.  
* `stratum` Character string specifying to which category of the stratum the timeseries belongs to when having added stratification parameters (more details in the next example). For unstratified timeseries it is `NA`.  
* `method` Character string specifying the signal detection method which was used. 
* `number_of_weeks` Integer specifying the number of weeks for which signals were generated.  



Generating stratified signals using EARS:
```{r}
data_prepro <- input_example %>% preprocess_data()
signals_ears_stratified <- data_prepro %>% get_signals(
  method = "ears",
  stratification = c("age_group", "county")
)
```

The `signals_ears_stratified` output is a data frame in long format, containing multiple time series of weekly aggregated case counts. Each time series corresponds to a unique combination of the specified stratification variables, with all strata stacked vertically.  

The columns `category` and `stratum` identify the individual timeseries. In this example we obtain the following values:  
* `category` is filled with the column names provided in the `stratification` parameter.  
* `stratum` with the values of these variables.  

In this example we obtain:  
* `category` is filled with "age_group" and "county"  
* `stratum` contains "00–04", "05–09", "10–14", "15–19", "20–24", "25–29", "30–34", "35–39", "40–44",
"45–49", "50–54", "55–59", "60–64", "65–69", "70–74", "75–79", "80–84", "85–89",
"90–94", "95–99", "100–104", "105–109" for rows with `category == "age_group"`.  For rows with `category == "county"` the values of `stratum` are "Burgenland", "Kärnten", "Niederösterreich", "Oberösterreich", "Salzburg", 
"Steiermark", "Tirol", "Vorarlberg", "Wien".

### Create visualisations and tables

To visualise signal detection in your data for the most recent `number_of_weeks`, you first need to preprocess your linelist and then apply `get_signals()`. To display aggregated signal detection results, first aggregate the signals. In the final step, apply a function to generate either a plot or a table. The plotting functions support both interactive and static outputs; by default, static plots are generated. Tables are returned as interactive DataTables by default, with options to convert them to data.frame or flextable formats.

The following example shows how to visualise the time series of detected signals using the default parameters of FarringtonFlexible without stratification in `plot_time_series()`:

```{r, warning=FALSE, fig.width=7}
data_prepro <- input_example %>% preprocess_data()
signals <- data_prepro %>% get_signals()
signals_time_series <- signals %>% plot_time_series()
signals_time_series
```

In the time series visualisation, the detection period is shaded in blue. The expectation is shown as a black line. The threshold is indicated by a blue line. Weekly case counts are displayed as bars. A signal is generated for a given week when the number of cases exceeds the threshold.

To create an interactive or a non-interactive map you have to use aggregated signals and an appropriate shapefile in a combined sf-object which is used in `plot_regional()`. This is an example for visualising the specified data as a non-interactive map while using "county" as `stratum` and the internal European shapefile (`nuts_shp`).

```{r, warning=FALSE, fig.width=7}
data_prepro <- input_example %>% preprocess_data()
signals <- data_prepro %>% get_signals(stratification = c("county"))
signals_agg <- signals %>% aggregate_signals(number_of_weeks = 6)
nuts_shp <- nuts_shp %>%
  sf::st_as_sf() %>%
  dplyr::filter(LEVL_CODE == 2 & CNTR_CODE == "AT") %>%
  dplyr::select(NUTS_NAME, geometry)
shape_with_signals <- nuts_shp %>%
  dplyr::inner_join(
    signals_agg,
    by = c("NUTS_NAME" = "stratum")
  ) %>%
  sf::st_as_sf()
signals_map <- plot_regional(
  shape_with_signals,
  signals_agg_unknown_region = data.frame(),
  interactive = FALSE,
  toggle_alarms = FALSE
)
signals_map
```

In the non-interactive map visualisation, the number of cases per region is represented by the shade of blue, with darker shades indicating higher case counts. Regions without signals are shown with a black border. If at least one signal occurs in a region, the border is coloured red and the number of signals is displayed as a label within the region.

To create an interactive or a non-interactive barplot you have to use aggregated signals in `plot_barchart()`. This is an example for visualising the specified data using "age group" as `stratum` and for the non-interactive plot using `plot_barchart()`:

```{r, warning=FALSE}
data_prepro <- input_example %>% preprocess_data()
signals <- data_prepro %>% get_signals(stratification = c("age_group"))
signals_agg <- signals %>% aggregate_signals(number_of_weeks = 6)
signals_barchart <- plot_barchart(
  signals_agg,
  interactive = FALSE
)
signals_barchart
```

In the non-interactive barplot visualisation, the number of cases is represented by the height of the bars. If no signal occurs in a given stratum, the bar has no border. If at least one signal is detected, the bar is outlined in red. For non-interactive plots, the number of signals is additionally displayed as a label above the bar.

To create an interactive or a non-interactive 100% stacked barplot showing a week-to-week representation of strata levels that had alarms you have to use signal data and specify the number of stratification levels in `plot_signals_per_week()`. This is an example for visualising the specified data using "county" as `stratum` with 9 levels for the non-interactive plot:

```{r, warning=FALSE}
data_prepro <- preprocess_data(input_example)
signals <- get_signals(
  data_prepro,
  stratification = "county"
)
n.strata <- 9
signals_week_barchart <- plot_signals_per_week(
  signals,
  n_strata = n.strata
)
signals_week_barchart
```

In the non-interactive 100% stacked barplot visualisation, the proportion of levels with alarms per week is represented by the height of the red segment of each bar. If no signal occurs in a given week, the bar is shown entirely in grey.

To create a signals detection results table with different formatting options you can use `build_signals_table()`. This is an example for creating a DataTable using age_group as `stratum` and default options: 

```{r, warning=FALSE, results='asis'}
data_prepro <- input_example %>% preprocess_data()
signals <- data_prepro %>% get_signals(
  stratification = c("age_group"),
  number_of_weeks = 6
)
signals_table <- signals %>% build_signals_table()
signals_table
```

To create an aggregated signal detection results table with different formatting options you can use `build_signals_agg_table()`. This is an example for creating a DataTable using age_group as `stratum` and default options: 

```{r, warning=FALSE, results='asis'}
data_prepro <- input_example %>% preprocess_data()
signals <- data_prepro %>% get_signals(
  stratification = c("age_group"),
  number_of_weeks = 6
)
agg_signals <- signals %>% aggregate_signals(number_of_weeks = 6)
agg_signals_table <- agg_signals %>% build_signals_agg_table()
agg_signals_table
```

In the table, rows with a signal are highlighted in red.

### Run the report
#### Single-pathogen reporting

You can directly generate an HTML or Word report containing signal results and visualizations using `run_report()`. Simply pass your surveillance linelist—structured according to the format defined in `input_metadata`—to `run_report()`. The data will be preprocessed automatically within the function. The following code chunk generates a Word report using EARS without any stratification (default) for the past six weeks (default):

```{r, eval=FALSE}
run_report(input_example,
  report_format = "DOCX",
  method = "EARS"
)
```

This generates an HTML report (default) with stratification by age group, county and sex for the last 2 weeks using FarringtonFlexible (default):

```{r, eval=FALSE}
run_report(input_example,
  strata = c("age_group", "county", "sex"),
  number_of_weeks = 2
)
```

#### Multi-pathogen reporting (HTML)

The following functionalities are currently only available for HTML reports. They are not supported in Word (DOCX) format.
The package also supports generating a single HTML report for multiple pathogens.  
To use this functionality, you must provide a line list that includes cases from different pathogens.

An example line list is available in the package as an internal dataset: `input_example_multipathogen`.

You can access it with:

```{r, eval=FALSE}
data("input_example_multipathogen")
```

To run an HTML report with the EARS algorithm and stratification by age_group and county for all pathogens in the line list you can run: 

```{r, eval=FALSE}
run_report(input_example_multipathogen,
  strata = c("age_group", "county"),
  method = "EARS"
)
```

If you want to generate a report for a subset of pathogens in your linelist you can specify these using the pathogens parameter:

```{r, eval=FALSE}
run_report(input_example_multipathogen,
  pathogens = c("Enterobacter", "Salmonella"),
  strata = c("age_group", "county"),
  method = "EARS"
)
```

In addition you can customise the HTML report by replacing the default United4Surveillance logo in the top right corner with your own logo. To do this, provide the path to a .png or .svg file using the custom_logo parameter.

```{r, eval=FALSE}
run_report(input_example_multipathogen,
  pathogens = c("Enterobacter", "Salmonella"),
  strata = c("age_group", "county"),
  method = "EARS",
  custom_logo = "path/to/my/custom_logo.png"
)
```
