---
title: "Signal Detection Report - `r params$country`"
date: "Report generated on `r Sys.Date()`"
lang: "en"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll

params: 
  data: NULL
  disease: "Pertussis"
  country: "Austria"
  number_of_weeks: 6
  method: "farrington"
  strata: !r c("county_id", "community_id", "sex", "age_group") 
  tables: true
  signals_padded: NULL
  signals_agg: NULL
  intervention_date: NULL

---


```{r, echo=FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      echo = FALSE, 
                      fig.align = 'center',
                      fig.width = 13,
                      fig.height = 5,
                      message = FALSE, 
                      warning = FALSE)
```

```{r include=FALSE}
# loading libraries in landing page level
library(dplyr)
library(SignalDetectionTool)
library(htmltools)
library(flexdashboard)
```

<!-- Landing page -->

Landing Page
================================================================================

```{r}
# For calculating week of start and end
records <- signals_padded %>% 
  dplyr::filter(!is.na(alarms)) %>% 
  dplyr::count(year, week) %>% 
  dplyr::mutate(yearweek = sprintf("%d-W%02d", year, week))

# Selected pathogens
text_diseases <- ifelse(
  length(params$disease) == 1, 
  params$disease, 
  paste0(paste0(params$disease, collapse = ", "),
         ", ")
)

# selected stratifications
text_strata <- ""
if(length(params$strata) >= 1){
  text_strata <- ", as well as stratified by "
  if(length(params$strata) == 1){
    text_strata <- paste0(text_strata, params$strata)
  } else if(length(params$strata) == 2){
    text_strata <- paste0(text_strata,
                          params$strata[1], " and ", params$strata[2])
  } else {
    text_strata <- paste0(text_strata,
                          paste0(params$strata[1:(length(params$strata) - 1)],
                                 collapse = ", "), 
                          ", and ", tail(params$strata, 1))
  }
} 
```


This report contains an overview of signals observed for the period `r head(records$yearweek, 1)` to `r tail(records$yearweek, 1)` for `r text_diseases` in `r params$country`. The signals were calculated using the `r params$method` algorithm.

For each disease, you can find an unstratified overview of the observed signals`r text_strata`. 

```{r}
results_cases <- params$signals_padded %>% 
  dplyr::filter(!is.na(alarms)) %>% 
  dplyr::group_by(pathogen, category, stratum) %>% 
  dplyr::summarise(cases = sum(cases),
                   signals = sum(alarms))
```

For the last `r params$number_of_weeks` weeks there were:
  
```{r, results='asis'}
for(pat in params$disease){
  pat_link <- paste0("[", pat, "]")
  tmp <- results_cases %>% dplyr::filter(pathogen == pat, is.na(category))
  cat("-", tmp$cases, "cases of", pat_link, "\n")
  cat("\t -", tmp$signals, "signals (unstratified) \n")
  for(cate in params$strata){
    tmp <- results_cases %>% dplyr::filter(pathogen == pat, category == cate)
    cat("\t -", sum(tmp$signals), "signals stratified by", cate, "\n")
  }
}
```


<!-- Pathogen pages -->

```{r, results='asis'}
src <- lapply(params$disease, function(i) {
  knitr::knit_expand('SignalDetectionReport_body.rmd', 
                     pathogen = i)
})

res <- knitr::knit_child(text = unlist(src), quiet = TRUE)
cat(res, sep = "\n")
```




