---
output: 
  html_document:
    toc: true
---

```{r, echo=FALSE}
knitr::opts_chunk$set(eval = TRUE, 
                      echo = FALSE, 
                      fig.align = 'center',
                      fig.width = 13,
                      fig.height = 5,
                      message = FALSE, 
                      warning = FALSE)
```

```{r libraries}
library(dplyr)
library(flextable)
```

# Introduction
This tab is meant to help, guide and assist you with any issues or questions that might arise while using the SignalDetectionTool app. It will cover the background and purpose of the tool, guide you through the different options and functionalities provided, and provides a contact point for reporting any bugs found while using the tool.

## United4Surveillance 
The SignalDetectionTool is developed under the United4Surveillance (U4S) project, work package 2, Task 2 “Outbreak and Signal Detection”. The goal of U4S is to ensure better detection of early warning signs and to strengthen infectious disease surveillance systems at the national level. Through early detection of infectious disease outbreaks, the further spread of diseases can potentially be contained, thereby reducing harmful consequences. Automated signal detection using tools can thereby help identify specific events of concern and manage and foster automated data analysis.

## Why use the SignalDetectionTool? 
Use this Shiny app as what it is - a tool. A tool to help bring another perspective to your disease surveillance, while also making it easy to generate reports and figures.

# Input data
This app requires specific input data formatting, which is described in detail in MS Teams.

# Workflow of the tool
The SignalDetectionTool app is made up of several tabs, each holding different key properties for using the app correctly. The workflow of the tool is to move from left to right, in a natural order. However, it may be necessary to jump between the `Input` and `Signals` tabs, to achieve the right signal detection for your needs. Additionally, the `Input`, `Signals` and `Report` tabs will be empty until an input data set is given.

## Data tab
The data tab is where you provide the tool with the proper input data to perform signal detection upon.

Clicking on the `Browse` button, you access a File Manager System, where you can navigate to your data set, saved as either an Excel (.xlsx) or comma-seperated file (.csv). 
Once you have selected a file, the file is read and checked by the tool. You will then receive feedback about your data in two sections: 

* In the first section, only those columns in your data where the column name has been written correctly are checked. For example, the column “country_id” is not checked when the column name deviates from it and is, for example, “country_id(NUTS)” in your dataset. This section checks whether all mandatory columns are present in your dataset and whether your columns have the right type and correct values. For example, it is checked whether the dates are in the correct YYYY-MM-DD format.

* In the second section, you receive feedback about all columns in your dataset that differ from the predefined names and are therefore not recognised by the DataCheckTool. From the example given above in i) the column “country_id(NUTS)” would appear in this section. As you want to use the column “country_id(NUTS)” later for signal detection, please change the column names to the required ones. Columns shown in this section are not checked for correct type and values.

In general, this feedback mechanism can, and should, be used in an iterative process. This means, that you get feedback about your data, you correct your data accordingly, and then upload it again. When all errors are fixed you can stop, otherwise you can continue correcting and uploading.
In the data tab, you are also presented with a table showing the data provided. If the input data did not pass the data format check, then the raw data provided is shown. Otherwise a table of pre-processed data, ready for signal detection is presented and you can move on the next tab,`Input`.

## Input tab 
The input tab is where you can interact, manipulate and shape the input data set, to investigate for signals proper. There are multiple different features to change, which will be covered in this section.

### Pathogen
Using this tool, it is only possible to perform signal detection upon one kind of pathogen at a time. Thus, if data about multiple pathogens exist in data, one will be preselected and can be changed using the drop-down list.

### Filters
You can choose to investigate a subset of your data according to the filters you select. When filtering upon one or more variables, whether it be choosing a specific time period through date_report and / or selecting only the age_groups 00-29, that subset of data is then used to train the outbreak detection algorithms and is shown on the timeseries visualisation.

### Strata
You can choose to stratify by up to three variables. Choosing a variable for stratification means that signals will be examined for each level of chosen strata(s). For example, if the variable age_group is one of chosen strata, then each age_group level, e.g. "00-04", "05-09" etc., will be examined for signals. This strata examination will be shown both in the result table found at the bottom of the `Signals` tab, and additionally, a strata-specific visualisation or table is shown on the top of the `Signals` tab. Choosing a strata will not affect the time series visualisation in the `Signals` tab.

### Signal detection period
Here you can set the number of weeks you want to generate signals for. The signals are generated for the most recent weeks, which is also dependent on if any filtering is done.

### Signal detection algorithm
Depending on the number of weeks you want to generate signals for, and the filters you set, the choice of algorithms is automatically updated to those which are possible to apply for your setting, with respect to the individual needs of training data of the different algorithms.
In the following, we call the training period the time period which is used to fit the parameters of the algorithms. This time period is given by your selection of date_report (Filters section) and the Signal detection period. If you select the time period 01-01-2020 until 31-12-2021 in the Filters section and select six weeks for the signal detection period the training period will be from 01-01-2020 until 22-11-2021 (six weeks before 31-12-2021). If you do not set any filter for date_report in the Filters section then the full time period of your data set minus the Signal Detection Period will be used to fit the parameters.

Which signal detection algorithm should you choose? It depends on multiple criteria, and they all have their individual weaknesses and strengths. The table below provides an overview of key characteristics, presented in the same order as the algorithms appear in the app's dropdown menu. A 'Yes' under 'Time trend' indicates that the selected signal detection method can model a time trend in the data. A 'Yes' under 'Seasonality' indicates that the signal detection method accounts for seasonal variations in case numbers. The minimum amount of historical data required varies by method. In particular, methods that model seasonality demand more historical data to accurately estimate seasonal patterns. As noted above, the app automatically determines which methods can be applied based on the amount of historical data available in your dataset and only shows you those.

```{r}
algorithms_overview <- readxl::read_excel(system.file("extdata", "help_algorithms.xlsx", package = "SignalDetectionTool"))
flextable(algorithms_overview) %>% 
  bold(part = "header") %>% 
  vline(part = "all") %>% 
  border_outer() %>% 
  autofit()
```

For more details about the individual algorithms we recommend to read the texts below. 

#### CUSUM  

CUSUM (cumulative sum) is an algorithm that detects shifts in the mean of a time series. It accumulates deviations of the observed case counts from the mean case count and once this accumulation has reached a value above a defined tolerance a signal is triggered. After a signal is triggered, the cumulative sum is set to zero, and the algorithm restarts. As CUSUM accumulates the deviations over time, it is more sensitive to small and moderate shifts in the case count. The mean case count is computed on all case counts of the training period. CUSUM does not account for seasonality as it just computes a mean over all case counts. The SignalDetectionTool calls the function cusum_with_reset(), an adaptation of the function algo.cusum() from the R package surveillance (Salmon et al, 2016) while using its default parameters.
CUSUM was originally developed by Page (1954) and typically used in industry for monitoring change detection. Rossi et al. (1999) serves as reference in the public health context.

#### FarringtonFlexible

The FarringtonFlexible uses a generalised linear model (GLM) to model the observed case count. In addition to a time trend terms for seasonality are also included in this model. For the computation of the baseline past very high case counts (potential outbreaks) are downweighted so that they don’t shift the baseline upwards. Furthermore, the past 26 time points before the current time point are excluded from the baseline to avoid reducing sensitivity when an outbreak has recently started before the current time point. The algorithm needs at least 78 time points (1.5 years when time points are weeks) of data in the past to calibrate the baseline.
FarringtonFlexible (Noufaily et al. 2013) is an improved version of the classic Farrington algorithm (Farrington et al. 1996). In this tool the implementation farringtonFlexible() from the R package surveillance (Salmon et al, 2016) with its default parameters is used. 

#### EARS 

The EARS (Early Aberration Reporting System) algorithm uses a window-based approach for outbreak detection. For the EARS C1 algorithm implemented in this tool, mean and standard deviation are computed from the cases observed in the last 7 time points before the current time point. These parameters are used to compute a threshold based on the quantiles of the normal distribution. When the observed case count is above the computed threshold a signal is raised. Because of the short time interval considered, EARS does not take into account seasonality and time trends. The baseline is computed using only the last 7 time points.
EARS was developed by CDC and published in Hutwagner et al. (2003). In this tool the implementation earsC() from the R package surveillance (Salmon et al, 2016) with its default parameters is used.  

#### Mean 

The algorithms labelled “Mean”, “Timetrend”, “Harmonic”, “Harmonic with timetrend”, “Step harmonic”, and “Step harmonic with
timetrend” share the same generalised linear model (GLM) backbone implemented in the function get_signals_glm(). Aggregated case counts are iteratively refitted with a quasi-Poisson GLM where down-weighted Anscombe residuals from reduce the influence of past outbreaks. Depending on the method, the GLM contains an intercept only, time trends, harmonic sine/cosine waves, or terms for 10 seasonal windows. A threshold is then determined using a prediction interval derived from the mean, applying either a Poisson or negative binomial distribution. The choice of distribution is based on the fitted dispersion of the data. 

The Mean option fits only an intercept, i.e., calculates the mean of all previous case counts while downweighting past outbreaks. This mean serves as the expected value for the current number of cases.

#### Timetrend

This method extends the intercept-only model (as used by “Mean”) by adding a linear timetrend.

#### Harmonic

This method augments the intercept-only model (as used by “Mean”) with sine and cosine regressors to model seasonality in the data. 

#### Harmonic with timetrend

The same as the Harmonic method but with fitting a timetrend to account for the underlying trend in the data.


#### Step harmonic

This method models the expected number of cases using the season groups defined in the Farrington outbreak detection algorithm. The year is divided into 10 seasonal windows, each with a constant expected case count within its respective season group. This approach offers flexibility in capturing seasonal patterns, such as a faster decline following a rise, but it requires more data for accurate fitting. 

#### Step harmonic with timetrend

The same as the Step harmonic method but with fitting a timetrend to account for the underlying trend in the data.

### Covid19 Pandemic correction

The "Covid19 Pandemic Correction" feature, accessible via an additional button, applies an interrupted time series analysis to the outbreak detection methods. It is possible to apply the correction to the following algorithms: Mean, Timetrend, Harmonic, Harmonic with timetrend, Step harmonic, Step harmonic with timetrend.  
In the Mean model, separate means are fitted for the periods before and after the selected intervention date. For models that include a time trend, both the intercept and time trend can differ following the intervention. However, the seasonality components are fitted across the entire time period and are not adjusted based on the intervention date.
This approach allows you to model changes in case counts, such as a decline due to the pandemic, and to capture a different time trend that may emerge afterward. A critical assumption behind these models is that the current case counts are represented by the model applied after the intervention, meaning it models the case counts which occured during and following the pandemic.

### References

- Farrington, C. P., Andrews, N. J., Beale, A. D., & Catchpole, M. A. (1996). A statistical algorithm for the early detection of outbreaks of infectious disease. *Journal of the Royal Statistical Society: Series A (Statistics in Society)*, 159(3), 547–563.
- Hutwagner, L., Thompson, W., Seeman, G. M., & Treadwell, T. (2003). The bioterrorism preparedness and response Early Aberration Reporting System (EARS). *Journal of Urban Health*, 80(2), i89–i96.
- Noufaily, A., Enki, D. G., Farrington, C. P., Garthwaite, P. H., Andrews, N., & Charlett, A. (2013). An improved algorithm for outbreak detection in multiple surveillance systems. *Statistics in Medicine*, 32(7), 1206–1222.
- Page, E. S. (1954). Continuous inspection schemes. *Biometrika*, 41(1/2), 100–115.
- Rossi, G., Lampugnani, L., & Marchi, M. (1999). An evaluation of statistical models for the early detection of infectious disease outbreaks. *Statistics in Medicine*, 18(23), 3323–3333.
- Salmon, M., Schumacher, D., & Höhle, M. (2016). Monitoring count time series in R: Aberration detection in public health surveillance. *Journal of Statistical Software*, 70(10), 1–35.


## Signals tab

The `Signals` tab is split into 2 / 3 "rows", depending on whether any strata were chosen in the `Input` tab. The first row belongs to the visualisations of the respective strata chosen. 
The second row to a time series / Epi Curve, depicting the signal detection in the number of weeks specified in the `Input` tab.
The third and final row display a signal detection table, which provides information on the signals found in both the entirety of the (filtered) data set as well as per stratification parameter.

In the time series visualisation the expectation and/or the threshold is also extended into the past for a better understanding of how the algorithm used works. For the algorithms Mean, Timetrend, Harmonic, Harmonic with timetrend, Step harmonic and Step harmonic with timetrend we display the expected number of cases from the model fitted to the first (earliest with respect to time) timepoint of the signal detection period shown blue. For the expected values displayed during the signal detection period, the model is refitted to a time series with additional historical data. As each new week of the signal detection period is included, more past weeks are incorporated into the fitting process, causing the mean expectation to change over time during the blue signal detection period.

### Stratification visualisation 
When one of the strata about regions (state, county or community) is selected in the `Input` tab the app tries to provide a map visualisation. This currently only works if there are NUTS codes provided in the data using the corresponding _id columns (i.e. county_id) and these NUTS ids all match the shapefile which is used in the app. When your data has cases with missing region during the signal detection timeperiod a map with additional text is shown which gives you information about these cases with missing/unknown region. In case you did not provide NUTS ids or some NUTS ids are not matching no map visualisation is currently possible and you will see a table with the results instead of a map. 

## Report tab
The `Report` tab is the final destination in the SignalDetectionTool Shiny app. Here, it is possible to produce and download a report, using the different parameter selections, filtering choices and stratifications chosen in the `Input` tab. Thus, you can customise exactly how you want your report. 
The report can be produced in two different formats: HTML or DOCX. 

# Bug reporting
If you encounter any technical issues or bugs when using the SignalDetectionTool, please report them on [Github](https://github.com/United4Surveillance/signal-detection-evaluation-tool/issues/new/choose) by making an issue. 
Thank you in advance!

